Correspondence Analysis of Two Mode Networks

### 2025-01-27 main-R1.tex

\subsection{Empirical Example} \label{subsec:ex}
Figure~\ref{fig:p-refs} shows the trajectory of the HH reflective centralities for persons in the Southern Women data \citep{davis1941}, using a bump chart to plot the rank order trajectory of persons across reflections. The rank order of people and groups in the $q^{th}$ centrality is plotted on the y-axis, and the centrality iteration $q$ is plotted on the x-axis. As the figure shows, $NORA$, $FLORA$, $CHARLOTTE$, and $EVELYN$ are the top-ranked actors when it comes to $C^R_p(2)$:  The average number of members of the groups they belong to. However, their fates in this reflective metric diverge at higher reflections, with $NORA$ and $FLORA$ maintaining their top positions but $CHARLOTTE$ and $EVELYN$ tumbling down the ranks, suggesting that the members of the groups they belong to affiliate with smaller groups than the members of the groups $NORA$ and $FLORA$ belong to (and so on for higher reflections). Notably, the reflective centrality rankings after freezing ($q \geq 18$) recover the Doreian et al. (2004) block partition (see Table~\ref{tab:southern}), but this time with Dorothy and Pearl in the middle separating the two largest blocks of women.

>>>>

\subsection{Empirical Example} \label{subsec:ex}
\subsubsection{Souther Women Data}
The following analyses use the classic Southern Women data \citep{davis1941} (DGG) as the running example, a dataset that, according to \citet[p.]{freeman2003finding}, ``has become something of a touchstone for comparing analytic methods in social network analysis,'' having been analyzed ``again and again,'' and reappearing ``whenever any network analyst wants to explore the utility of some new tool for analyzing data.'' Freeman himself would go to review twenty-one such attempts, and the list has only gotten longer since his analytic ``meta-analysis'' \citep[e.g.,][]{doreian2004generalized, field2006identifying, roffilli2006identifying, wang2009exponential, everett2018measuring, lerner2022dynamic, batagelj2022analysis,  lizardo2024two}. As we can see, novel insights can still be garnered from these old data.

One key source of confusion in the mini-literature that exists around this dataset concern the order of presentation of the rows and columns. The reason for this is while the row mode (people) is ``nominal'' in both the scaling and the literal senses (being composed of women's names), the column mode (events) is more ambiguous. It \textit{could} be treated as nominal, but the original DGG table includes event \textit{dates}, which means that it could also be treated as an ordered variable.  As \citet{freeman2003finding} notes, in the canonical original table both the row and column modes had be re-ordered in presentation to reveal what DGG thought was the group structure among the women; this means that the events were \textit{not} included in chronological order. To add to the confusion, the original DGG table added arbitrary ordered numerical designators to both people and events, such that people came to refer to "woman 1" or woman 11" and "E1" or "E13" based on the original DGG ordering, which may not have mattered for the women, but it mattered for the events because "E14" in the original DGG data is not the last chronological even among the fourteen columns, but in fact, as shown in~\ref{tab:southern}, it's the event that occurred on August 3rd, and there were three events that occurred after that one (E4, E8, and E13) in the DGG numbering.

To further add to the confusion, some early analyses of the data, including Breiger's \citeyearpar{breiger1974duality} classic article and \citet{doreian1979evolution} used a \textit{different} ordering of the women (taken from \citet{homans}) from that of the original DGG table reproduced in \citet{freeman2003finding}, with the events ordered chronologically. This means that the arbitrary ordinal labeling of the rows (e.g., "w1, "w2", "w3", etc.) and the columns (e.g., "E1, "E2", "E3", etc.) in these publications---see e.g., \citep[table 1]{doreian1979evolution}---does not correspond to the DGG ordering reproduced in \citet{freeman2003finding}. Different versions of the dataset reproduce different versions of the labeling. Of course, at the end of the day, the specific labeling does not really matter, since we are only interested in the structural signature of the rows and columns. The labeling does matter, however, when trying to reconcile claims across publications from the narrative, since "Woman 1" or "E8" in one publication (e.g., \citet{freeman2003finding}) may not correspond to correspond to "Woman 1" or "E8" in the other. To forestall confusion, in what follows I use the numeric labeling corresponding to the original DGG data table as reproduced in \citet[Figure 1]{freeman2003finding}, and shown in Table~\ref{tab:southern}, which include both the arbitrary ordered labels (e.g., "w1" or "E1") and both the woman's name and the event date,, with "Myrna" rebaptized to "Myra" to correspond to the original DGG table \citep{freeman2003finding}.

\subsubsection{Reflective Score Trajectory}

Figure~\ref{fig:p-refs} shows the trajectory of the HH reflective centralities for persons, using a bump chart to plot the rank order trajectory of persons across reflections. The rank order of people and groups in the $q^{th}$ centrality is plotted on the y-axis, and the centrality iteration $q$ is plotted on the x-axis. As the figure shows, $NORA$, $FLORA$, $CHARLOTTE$, and $EVELYN$ are the top-ranked actors when it comes to $C^R_p(2)$:  The average number of members of the groups they belong to. However, their fates in this reflective metric diverge at higher reflections, with $NORA$ and $FLORA$ maintaining their top positions but $CHARLOTTE$ and $EVELYN$ tumbling down the ranks, suggesting that the members of the groups they belong to affiliate with smaller groups than the members of the groups $NORA$ and $FLORA$ belong to (and so on for higher reflections). Notably, the reflective centrality rankings after freezing ($q \geq 18$) recover the Doreian et al. (2004) block partition (see Table~\ref{tab:southern}), but this time with Dorothy and Pearl in the middle separating the two largest blocks of women.
